Proofs for ``Geometry of Local Spectral Clustering''

### 2021-01-20 jmlr/revision/proofs.tex
Completed revised proof of lower bound.

The proofs of our major theorems all consist of at most three modular parts.
\begin{enumerate}
	\item~\textbf{Fixed graph results.} Results which hold with respect to an arbitrary graph $G$, and are stated with respect to functionals (i.e. normalized cut, conductance, and local spread) of $G$;
	\item~\textbf{Finite-sample bounds.} For the specific choice of $G = G_{n,r}$, we relate the aforementioned functionals to their population analogues. This proves~\textcolor{red}{(?)} 
	\item~\textbf{Population functionals.} (In the case of density clustering only.) When the candidate cluster is a $\lambda$-density cluster, we bound the population functionals by a function of $\lambda$, as well as the other relevant parameters introduced in Section~\textcolor{red}{(?)}. This proves~\textcolor{red}{(?)}
\end{enumerate}
The first three sections of this appendix will correspond to each of these three parts. In Section~\ref{sec:pf_major_theorems}, we will combine these parts to prove the major theorems of our main text, \textcolor{red}{(...)}. Finally, in Section~\ref{sec:appr_misclassification_error} we derive upper bounds for the aPPR vector, and in Section~\ref{sec:experimental_details} we give relevant details regarding our experiments.

@@

\subsection{Sample-to-population: local spread}
\label{subsec:sample_to_population_local_spread}
In this subsection we establish~\eqref{eqn:sample_to_population_local_spread}. To ease the notational burden, let $\wt{G}_{n,r} := G_{n,r}\bigl[\mc{C}[X]\bigr]$. By Lemma~\ref{lem:hoeffding}, with probability at least $1 - \exp\{-2\delta^2 \Pbb(\mc{C})^2n\}$
\begin{equation}
\label{pf:sample_to_population_local_spread_1}
\wt{n} \geq (1 - \delta) \cdot n \cdot \Pbb(\mc{C}).
\end{equation}
Conditional on~\eqref{pf:sample_to_population_local_spread_1}, it follows from Lemma~\ref{lem:bernstein_union} that with probability at least $1 -n\exp\{-2(1 - \delta)\delta^2d_{\min}(\wt{\Pbb})^2 \Pbb(\mc{C}) n\}$,
\begin{equation}
\label{pf:sample_to_population_local_spread_2}
\frac{1}{\wt{n}} d_{\min}(\wt{G}_{n,r}) \geq (1 - \delta) \cdot d_{\min}(\wt{\Pbb}),
\end{equation}
and likewise it follows from Lemma~\ref{lem:hoeffding_2} that with probability at least $1 - \exp\{-2\delta^2(1 - \delta) \vol_{\wt{\Pbb},r}(\mc{C})^2 \Pbb(\mc{C})n \}$,
\begin{equation*}
\frac{1}{\wt{n}(\wt{n} - 1)} \vol(\wt{G}_{n,r}) \geq (1 - \delta) \cdot \vol_{\wt{\Pbb},r}(\mc{C}).
\end{equation*}
Consequently, for any $\delta \in (0,1/3)$,
\begin{equation*}
s_{n,r}(\mc{C}[X]) = \frac{d_{\min}(\wt{G}_{n,r})^2}{\vol(\wt{G}_{n,r})} \geq \frac{\frac{1}{\wt{n}^2}d_{\min}(\wt{G}_{n,r})^2}{\frac{1}{\wt{n}(\wt{n} - 1)}\vol(\wt{G}_{n,r})} \geq \frac{(1 - \delta)^2}{(1 + \delta)} \frac{d_{\min}(\wt{\Pbb})^2}{\vol_{\wt{\Pbb},r}(\mc{C})} \geq (1 - 3\delta) \cdot s_{\Pbb,r}(\mc{C}).
\end{equation*}
with probability at least $1 - 2\exp\bigl\{-\delta^2\bigl(\vol_{\wt{\Pbb},r}(\mc{C})\bigr)^2 \Pbb(\mc{C})n\bigr\} - n\exp\bigl\{-\delta^2\bigl(d_{\min}(\wt{\Pbb})\bigr)^2 \Pbb(\mc{C}) n\bigr\}$. This establishes~\eqref{eqn:sample_to_population_local_spread} upon taking $b_2 = \bigl(\vol_{\wt{\Pbb},r}(\mc{C})\bigr)^2 \Pbb(\mc{C})$, $b_3 = \bigl(d_{\min}(\wt{\Pbb})\bigr)^2 \Pbb(\mc{C})$, and $C_2 = 3$.

@@

\paragraph{Putting the pieces together.} 
First, we note that
\begin{equation*}
\Psi_r(\wt{\mbb{P}}) = \Psi_{\wt{\mbb{P}},r}(\mathrm{supp}(\wt{\mbb{P}})) = \Psi_{\Pbb,r}(\mc{C}), ~~\textrm{and}~~\Psi_{r}(\wt{\mbb{P}}_n) = \Psi_{\wt{\mbb{P}}_n, r}(\mathrm{supp}(\wt{\mbb{P}}_n)) = \Psi_{n,r}(\mc{C}[X]),
\end{equation*}
so that we may apply Proposition~\ref{prop:conductance_lb_transportation_distance} to get a lower bound on $\Psi_{n,r}(\mc{C}[X])$ in terms of $\Psi_{\Pbb,r}(\mc{C})$, $\Delta_{\infty}(\wt{\Pbb}, \wt{\Pbb}_n)$, and $d_{\min}(\wt{\Pbb}_{n})$. Next, as we have already derived in~\eqref{pf:sample_to_population_local_spread_2}, with probability at least $1 - \exp\{-\mbb{P}(\mc{C})^2n\} - \exp\{-d_{\min}(\wt{\Pbb})^2 \Pbb(\mc{C}) n/\sqrt{2}\}$,
\begin{equation*}
d_{\min}(\wt{\Pbb}_n) = \frac{1}{\wt{n}} d_{\min}(\wt{G}_{n,r}) \geq \frac{1}{\sqrt{2}} d_{\min}(\wt{\Pbb}).
\end{equation*}
Finally, taking
\begin{equation*}
b_5 := \frac{1}{2b_4}\Psi_r(\wt{\Pbb}) \cdot (1 - \Psi_r(\wt{\Pbb})) \cdot d_{\min}(\wt{\Pbb})^2,~~\textrm{and}~~ B_1 := \frac{1}{B_4}\min\Bigl\{\frac{b_5}{2\Psi_r(\wt{\Pbb})}, b_5, \frac{r}{4(d - 1)} \Bigr\} 
\end{equation*}
by Theorem~\ref{thm:garciatrillos16} and~\eqref{eqn:sample_to_population_conductance_sample_complexity} we have that
\begin{equation*}
\Delta_{\infty}(\wt{\Pbb},\wt{\Pbb}_n) \leq B_3 \frac{\log(n)^{p_d}}{n^{1/d}} \leq \min\Bigl\{\frac{b_5}{2\Psi_r(\wt{\Pbb})}, b_5, \frac{r}{4(d - 1)} \Bigr\} \cdot \delta
\end{equation*}
with probability at least $1 - B_2/n$. Therefore by Proposition~\ref{prop:conductance_lb_transportation_distance},
\begin{equation*}
\Psi_r(\wt{\Pbb}_n) \geq \Psi_r(\wt{\Pbb}) \cdot \biggl(1 - \frac{2b_4\Delta_{\infty}(\wt{\Pbb}_n,\wt{\Pbb})}{\bigl(1 - \Psi_r(\wt{\Pbb})\bigr) \cdot \bigl(d_{\min}(\wt{\Pbb}_n)\bigr)^2}\biggr) - \frac{b_4 \Delta_{\infty}(\wt{\Pbb}_n,\wt{\Pbb}_N)}{\bigl(1 - \Psi_r(\wt{\Pbb})\bigr) \cdot \bigl(d_{\min}(\wt{\Pbb}_n)\bigr)^2} \geq \Psi_r(\wt{\Pbb}) (1 - 2\delta)
\end{equation*}
with probability at least $1 - B_2/n - (n + 2)\exp\{-d_{\min}(\wt{\Pbb})^2 \Pbb(\mc{C}) n/\sqrt{2}\}$, establishing~\eqref{eqn:sample_to_population_conductance}.

\section{Population functionals for density clusters}
\label{sec:density_cluster_population_functionals}
In this section, we prove Lemma~\ref{lem:density_cluster_local_spread} (in Section~\ref{subsec:density_cluster_local_spread}), Proposition~\ref{prop:density_cluster_normalized_cut} (in Section~\ref{subsec:density_cluster_ncut}), and Proposition~\ref{prop:density_cluster_conductance} (in Section~\ref{subsec:density_cluster_conductance}), by establishing bounds on the population-level local spread, normalized cut, and conductance of a thickened density cluster $\mc{C}_{\lambda,\sigma}$. In these proofs, we make use of some estimates on the volume of spherical caps (given in Section~\ref{subsec:spherical_caps}); some isoperimetric inequalities (given in Section~\ref{subsec:isoperimetric_inequalities}), and some reverse isoperimetric inequalities (given in Section~\ref{subsec:reverse_isoperimetric_inequalities}). Throughout, we write $\nu_d := \nu(B(0,1))$ for the Lebesgue measure of a $d$-dimensional unit ball.

@@

For any set $\mc{C} \subseteq \Rd$ and $\sigma > 0$, recall that $\mc{C}_{\sigma} := \{x: \mathrm{dist}(x,\mc{C}) \leq \sigma\}$. We begin with an upper bound on the volume of $\mc{C}_{\sigma + \delta}$ as compared to $\mc{C}_{\sigma}$. 
\begin{lemma}
	\label{lem:reverse_isoperimetric_inequality}
	For any bounded set $\mc{C} \subseteq \Rd$ and  $\sigma, \delta > 0$, it holds that
	\begin{equation}
	\label{eqn:reverse_isoperimetric_inequality}
	\nu(\mc{C}_{\sigma + \delta}) \leq \nu(\mc{C}_{\sigma}) \cdot \Bigl(1 + \frac{\delta}{\sigma}\Bigr)^d.
	\end{equation}
\end{lemma}
Lemma~\ref{lem:reverse_isoperimetric_inequality} should be interpreted as a reverse isoperimetric inequality. To see this, note that if $\delta \leq \sigma/d$ then $\bigl(1 + \delta/\sigma\bigr)^d \leq 1 + d \cdot \delta/(\sigma - d\delta)$, and we deduce from~\eqref{eqn:reverse_isoperimetric_inequality} that
\begin{equation}
\label{eqn:reverse_isoperimetric_inequality_2}
\nu(\mc{C}_{\sigma + \delta} \setminus \mc{C}_{\sigma}) = \nu(\mc{C}_{\sigma + \delta}) - \nu(\mc{C}_{\delta}) \leq \frac{d\delta}{\sigma - d\delta} \cdot \nu(\mc{C}_{\sigma}).
\end{equation}
We use~\eqref{eqn:reverse_isoperimetric_inequality_2} along with Assumption~\ref{asmp:low_noise_density} to derive a density-weighted reverse isoperimetric inequality. 
\begin{lemma}
	\label{lem:reverse_isoperimetric_inequality_density_weighted}
	Let $\mc{C}_{\lambda,\sigma}$ satisfy Assumption~\ref{asmp:bounded_density} and~\ref{asmp:low_noise_density} for some $\theta, \gamma$ and $\lambda_{\sigma}$. Then for any $0 < r \leq \sigma/d$, it holds that
	\begin{equation}
	\label{eqn:reverse_isoperimetric_inequality_density_weighted}
	\Pbb\bigl(\mc{C}_{\lambda,\sigma + r} \setminus \mc{C}_{\lambda,\sigma}\bigr) \leq \Bigl(1 + \frac{dr}{\sigma - dr}\Bigr) \cdot \frac{dr}{\sigma} \cdot \left(\lambda_{\sigma} - \theta\frac{r^{\gamma}}{\gamma + 1}\right) \cdot \nu(\mc{C}_{\lambda,\sigma}).
	\end{equation}
\end{lemma}
\begin{proof}[Proof (of Lemma~\ref{lem:reverse_isoperimetric_inequality})] 
	Fix $\delta' > 0$, and take $\epsilon = \delta + \delta'$. We will show that
	\begin{equation*}
	\label{pf:reverse_isoperimetric_inequality}
	\nu(\mc{C}_{\sigma + \epsilon}) \leq \nu(\mc{C}_{\sigma}) \cdot \Bigl(1 + \frac{\epsilon}{\sigma}\Bigr)^d,
	\end{equation*}
	whence taking a limit as $\delta' \to 0$ yields the claim.

	To show~\eqref{pf:reverse_isoperimetric_inequality} for general closed and bounded $\mc{C}$, we need to construct a particular disjoint covering $\mc{A}_1(\sigma + \epsilon), \ldots, \mc{A}_N(\sigma + \epsilon)$ of $\mc{C}_{\sigma + \delta}$. To do so, we first take a finite set of points $x_1,\ldots,x_N$ such that the net $B(x_1,\sigma + \epsilon),\ldots,B(x_N,\sigma + \epsilon)$ covers $\mc{C}_{\sigma + \delta}$.  Note that such a covering exists for some finite $N = N(\epsilon)$ because $\mc{C}_{\sigma + \delta}$ is bounded, and the closure of $\mc{C}_{\sigma + \delta}$ is thus a compact subset of $\cup_{x \in \mc{C}} B(x,\sigma + \epsilon)$. Defining $\mc{A}_1(s), \ldots, \mc{A}_N(s)$ for a given $s > 0$ to be
	\begin{equation*}
	\mc{A}_1(s) := B(x_1,s),~~\textrm{and}~~ \mc{A}_{j + 1}(s) := B(x_{j + 1},s) \setminus \bigcup_{i = 1}^{j} B(x_i,s) ~~\textrm{for $j = 1,\ldots,N - 1$},
	\end{equation*}
	we have that $\mc{A}_1(\sigma + \epsilon),\ldots,\mc{A}_N(\sigma + \epsilon)$ is a disjoint covering of $\mc{C}_{\sigma + \delta}$, and so $\nu(\mc{C}_{\sigma + \delta}) \leq \sum_{j = 1}^{N} \nu(\mc{A}_j(\sigma + \epsilon))$.

@@

\subsection{Proof of Proposition~\ref{prop:density_cluster_conductance}}
\label{subsec:density_cluster_conductance}
The following Lemma upper bounds the population-level uniform conductance $\Psi_{\nu,r}(\mc{C}_{\lambda,\sigma})$. 
\begin{lemma}
	\label{lem:uniform_density_cluster_conductance}
	Suppose $\mc{C}_{\lambda,\sigma}$ satisfies Assumption~\ref{asmp:embedding} with respect to some $\rho > 0$ and $L \geq 1$. For any $0 < r \leq \sigma \sqrt{2\pi/(d + 2)}$, it holds that
	\begin{equation}
	\label{eqn:uniform_density_cluster_conductance}
	\Psi_{\nu,r}(\mc{C}_{\lambda,\sigma}) \geq \Bigl(1 - \frac{r}{4\rho L}\Bigr) \cdot \Bigl(1 - \frac{r}{\sigma}\sqrt{\frac{d + 2}{2\pi}}\Bigr)^2 \cdot \frac{\sqrt{2\pi}}{36} \cdot \frac{r}{\rho L \sqrt{d + 2}}.
	\end{equation}
\end{lemma}
Noting that $\Psi_{\Pbb,r}(\mc{C}_{\lambda,\sigma}) \geq \Psi_{\nu,r}(\mc{C}_{\lambda,\sigma}) \cdot \lambda_{\sigma}^2/\Lambda_{\sigma}^2$, Proposition~\ref{prop:density_cluster_conductance} follows from~\eqref{eqn:uniform_density_cluster_conductance}.

@@

@@

>>>>

The proofs of our major theorems largely consist of (at most) three modular parts.
\begin{enumerate}
	\item \textbf{Fixed graph results.} Results which hold with respect to an arbitrary graph $G$, and are stated with respect to functionals (i.e. normalized cut, conductance, and local spread) of $G$;
	\item \textbf{Finite-sample bounds.} For the specific choice of $G = G_{n,r}$, we relate the aforementioned functionals to their population analogues. This proves~\textcolor{red}{(?)} 
	\item \textbf{Population functionals.} (In the case of density clustering only.) When the candidate cluster is a $\lambda$-density cluster, we bound the population functionals by a function of $\lambda$, as well as the other relevant parameters introduced in Section~\textcolor{red}{(?)}. This proves~\textcolor{red}{(?)}
\end{enumerate}
The first three sections of this appendix will correspond to each of these three parts. In Section~\ref{sec:pf_major_theorems}, we will combine these parts to prove the major theorems of our main text, \textcolor{red}{(...)}. Finally, in Section~\ref{sec:appr_misclassification_error} we derive upper bounds for the aPPR vector, and in Section~\ref{sec:experimental_details} we give relevant details regarding our experiments.

@@

\subsection{Sample-to-population: local spread}
\label{subsec:sample_to_population_local_spread}
In this subsection we establish~\eqref{eqn:sample_to_population_local_spread}. To ease the notational burden, let $\wt{G}_{n,r} := G_{n,r}\bigl[\mc{C}[X]\bigr]$. By Lemma~\ref{lem:hoeffding}, with probability at least $1 - \exp\{-2\delta^2 \Pbb(\mc{C})^2n\}$
\begin{equation}
\label{pf:sample_to_population_local_spread_1}
\wt{n} \geq (1 - \delta) \cdot n \cdot \Pbb(\mc{C}).
\end{equation}
Conditional on~\eqref{pf:sample_to_population_local_spread_1}, it follows from Lemma~\ref{lem:bernstein_union} that with probability at least $1 -n\exp\{-2(1 - \delta)\delta^2d_{\min}(\wt{\Pbb}) \Pbb(\mc{C}) n\}$,
\begin{equation}
\label{pf:sample_to_population_local_spread_2}
\frac{1}{\wt{n}} d_{\min}(\wt{G}_{n,r}) \geq (1 - \delta) \cdot d_{\min}(\wt{\Pbb}),
\end{equation}
and likewise it follows from Lemma~\ref{lem:hoeffding_2} that with probability at least $1 - \exp\{-2\delta^2(1 - \delta) \vol_{\wt{\Pbb},r}(\mc{C})^2 \Pbb(\mc{C})n \}$,
\begin{equation*}
\frac{1}{\wt{n}(\wt{n} - 1)} \vol(\wt{G}_{n,r}) \geq (1 - \delta) \cdot \vol_{\wt{\Pbb},r}(\mc{C}).
\end{equation*}
Consequently, for any $\delta \in (0,1/3)$,
\begin{equation*}
s_{n,r}(\mc{C}[X]) = \frac{d_{\min}(\wt{G}_{n,r})^2}{\vol(\wt{G}_{n,r})} \geq \frac{\frac{1}{\wt{n}^2}d_{\min}(\wt{G}_{n,r})^2}{\frac{1}{\wt{n}(\wt{n} - 1)}\vol(\wt{G}_{n,r})} \geq \frac{(1 - \delta)^2}{(1 + \delta)} \frac{d_{\min}(\wt{\Pbb})^2}{\vol_{\wt{\Pbb},r}(\mc{C})} \geq (1 - 3\delta) \cdot s_{\Pbb,r}(\mc{C}).
\end{equation*}
with probability at least $1 - 2\exp\bigl\{-\delta^2\bigl(\vol_{\wt{\Pbb},r}(\mc{C})\bigr)^2 \Pbb(\mc{C})n\bigr\} - n\exp\bigl\{-\delta^2\bigl(d_{\min}(\wt{\Pbb})\bigr)^2 \Pbb(\mc{C}) n\bigr\}$. This establishes~\eqref{eqn:sample_to_population_local_spread} upon taking $b_2 = \bigl(\vol_{\wt{\Pbb},r}(\mc{C})\bigr)^2 \Pbb(\mc{C})$ and $b_3 = d_{\min}(\wt{\Pbb}) \Pbb(\mc{C})$

@@

\paragraph{Putting the pieces together.} 
First, we note that
\begin{equation*}
\Psi_r(\wt{\mbb{P}}) = \Psi_{\wt{\mbb{P}},r}(\mathrm{supp}(\wt{\mbb{P}})) = \Psi_{\Pbb,r}(\mc{C}), ~~\textrm{and}~~\Psi_{r}(\wt{\mbb{P}}_n) = \Psi_{\wt{\mbb{P}}_n, r}(\mathrm{supp}(\wt{\mbb{P}}_n)) = \Psi_{n,r}(\mc{C}[X]),
\end{equation*}
so that we may apply Proposition~\ref{prop:conductance_lb_transportation_distance} to get a lower bound on $\Psi_{n,r}(\mc{C}[X])$ in terms of $\Psi_{\Pbb,r}(\mc{C})$, $\Delta_{\infty}(\wt{\Pbb}, \wt{\Pbb}_n)$, and $d_{\min}(\wt{\Pbb}_{n})$. Next, as we have already derived in~\eqref{pf:sample_to_population_local_spread_2}, with probability at least $1 - \exp\{-\mbb{P}(\mc{C})n\} - \exp\{-d_{\min}(\wt{\Pbb})^2 \Pbb(\mc{C}) n/\sqrt{2}\}$,
\begin{equation*}
d_{\min}(\wt{\Pbb}_n) = \frac{1}{\wt{n}} d_{\min}(\wt{G}_{n,r}) \geq \frac{1}{\sqrt{2}} d_{\min}(\wt{\Pbb}).
\end{equation*}
Finally, taking
\begin{equation*}
b_5 := \frac{1}{2b_4}\Psi_r(\wt{\Pbb}) \cdot (1 - \Psi_r(\wt{\Pbb})) \cdot d_{\min}(\wt{\Pbb})^2,~~\textrm{and}~~ B_1 := \frac{1}{B_4}\min\Bigl\{\frac{b_5}{2\Psi_r(\wt{\Pbb})}, b_5, \frac{r}{4(d - 1)} \Bigr\} 
\end{equation*}
by Theorem~\ref{thm:garciatrillos16} and~\eqref{eqn:sample_to_population_conductance_sample_complexity} we have that
\begin{equation*}
\Delta_{\infty}(\wt{\Pbb},\wt{\Pbb}_n) \leq B_3 \frac{\log(n)^{p_d}}{n^{1/d}} \leq \min\Bigl\{\frac{b_5}{2\Psi_r(\wt{\Pbb})}, b_5, \frac{r}{4(d - 1)} \Bigr\} \cdot \delta
\end{equation*}
with probability at least $1 - B_2/n$. Therefore by Proposition~\ref{prop:conductance_lb_transportation_distance},
\begin{equation*}
\Psi_r(\wt{\Pbb}_n) \geq \Psi_r(\wt{\Pbb}) \cdot \biggl(1 - \frac{2b_4\Delta_{\infty}(\wt{\Pbb}_n,\wt{\Pbb})}{\bigl(1 - \Psi_r(\wt{\Pbb})\bigr) \cdot \bigl(d_{\min}(\wt{\Pbb}_n)\bigr)^2}\biggr) - \frac{b_4 \Delta_{\infty}(\wt{\Pbb}_n,\wt{\Pbb}_N)}{\bigl(1 - \Psi_r(\wt{\Pbb})\bigr) \cdot \bigl(d_{\min}(\wt{\Pbb}_n)\bigr)^2} \geq \Psi_r(\wt{\Pbb}) (1 - 2\delta)
\end{equation*}
with probability at least $1 - B_2/n - (n + 2)\exp\{-d_{\min}(\wt{\Pbb})\cdot \Pbb(\mc{C}) n/\sqrt{2}\}$, establishing~\eqref{eqn:sample_to_population_conductance}.

\section{Population functionals for density clusters}
\label{sec:density_cluster_population_functionals}
In this section, we prove Lemma~\ref{lem:density_cluster_local_spread} (in Section~\ref{subsec:density_cluster_local_spread}), Proposition~\ref{prop:density_cluster_normalized_cut} (in Section~\ref{subsec:density_cluster_ncut}), and Proposition~\ref{prop:density_cluster_conductance} (in Section~\ref{subsec:density_cluster_conductance}), by establishing bounds on the population-level local spread, normalized cut, and conductance of a thickened density cluster $\mc{C}_{\lambda,\sigma}$. In these proofs, we make use of some estimates on the volume of spherical caps (in Section~\ref{subsec:spherical_caps}); some isoperimetric inequalities (in Section~\ref{subsec:isoperimetric_inequalities}), and some reverse isoperimetric inequalities (given in Section~\ref{subsec:reverse_isoperimetric_inequalities}). Finally, in Section~\ref{subsec:density_cluster_hard_case}, for the hard case distribution $\Pbb$ defined in~\eqref{eqn:lb_density} and $\mc{L}$ defined in~\eqref{eqn:lower_set}, establish bounds on the population-level normalized cut $\Phi_{\Pbb,r}(\mc{L})$ and local spread $s_{\Pbb,r}(\mc{X})$; these will be useful in the proof of Theorem~\ref{thm:ppr_lb}. Throughout, we write $\nu_d := \nu(B(0,1))$ for the Lebesgue measure of a $d$-dimensional unit ball.

@@

For any set $\mc{C} \subseteq \Rd$ and $\sigma > 0$, recall that $\mc{C}_{\sigma} := \{x: \mathrm{dist}(x,\mc{C}) \leq \sigma\}$. We begin with an upper bound on the volume of $\mc{C}_{\sigma + \delta}$ as compared to $\mc{C}_{\sigma}$. 
\begin{lemma}
	\label{lem:reverse_isoperimetric_inequality}
	For any bounded set $\mc{C} \subseteq \Rd$ and  $\sigma, \delta > 0$, it holds that
	\begin{equation}
	\label{eqn:reverse_isoperimetric_inequality}
	\nu(\mc{C}_{\sigma + \delta}) \leq \nu(\mc{C}_{\sigma}) \cdot \Bigl(1 + \frac{\delta}{\sigma}\Bigr)^d.
	\end{equation}
\end{lemma}
Lemma~\ref{lem:reverse_isoperimetric_inequality} should be interpreted as a reverse isoperimetric inequality. To see this, note that if $\delta \leq \sigma/d$ then $\bigl(1 + \delta/\sigma\bigr)^d \leq 1 + d \cdot \delta/(\sigma - d\delta)$, and we deduce from~\eqref{eqn:reverse_isoperimetric_inequality} that
\begin{equation}
\label{eqn:reverse_isoperimetric_inequality_2}
\nu(\mc{C}_{\sigma + \delta} \setminus \mc{C}_{\sigma}) = \nu(\mc{C}_{\sigma + \delta}) - \nu(\mc{C}_{\delta}) \leq \frac{d\delta}{\sigma - d\delta} \cdot \nu(\mc{C}_{\sigma}).
\end{equation}
We use~\eqref{eqn:reverse_isoperimetric_inequality_2} along with Assumption~\ref{asmp:low_noise_density} to derive a density-weighted reverse isoperimetric inequality. 
\begin{lemma}
	\label{lem:reverse_isoperimetric_inequality_density_weighted}
	Let $\mc{C}_{\lambda,\sigma}$ satisfy Assumption~\ref{asmp:bounded_density} and~\ref{asmp:low_noise_density} for some $\theta, \gamma$ and $\lambda_{\sigma}$. Then for any $0 < r \leq \sigma/d$, it holds that
	\begin{equation}
	\label{eqn:reverse_isoperimetric_inequality_density_weighted}
	\Pbb\bigl(\mc{C}_{\lambda,\sigma + r} \setminus \mc{C}_{\lambda,\sigma}\bigr) \leq \Bigl(1 + \frac{dr}{\sigma - dr}\Bigr) \cdot \frac{dr}{\sigma} \cdot \left(\lambda_{\sigma} - \theta\frac{r^{\gamma}}{\gamma + 1}\right) \cdot \nu(\mc{C}_{\lambda,\sigma}).
	\end{equation}
\end{lemma}
\begin{proof}[Proof (of Lemma~\ref{lem:reverse_isoperimetric_inequality})] 
	Fix $\delta' > 0$, and take $\epsilon = \delta + \delta'$. We will show that
	\begin{equation}
	\label{pf:reverse_isoperimetric_inequality}
	\nu(\mc{C}_{\sigma + \epsilon}) \leq \nu(\mc{C}_{\sigma}) \cdot \Bigl(1 + \frac{\epsilon}{\sigma}\Bigr)^d,
	\end{equation}
	whence taking a limit as $\delta' \to 0$ yields the claim.

	To show~\eqref{pf:reverse_isoperimetric_inequality}, we need to construct a particular disjoint covering $\mc{A}_1(\sigma + \epsilon), \ldots, \mc{A}_N(\sigma + \epsilon)$ of $\mc{C}_{\sigma + \delta}$. To do so, we first take a finite set of points $x_1,\ldots,x_N$ such that the net $B(x_1,\sigma + \epsilon),\ldots,B(x_N,\sigma + \epsilon)$ covers $\mc{C}_{\sigma + \delta}$.  Note that such a covering exists for some finite $N = N(\epsilon)$ because $\mc{C}_{\sigma + \delta}$ is bounded, and the closure of $\mc{C}_{\sigma + \delta}$ is thus a compact subset of $\cup_{x \in \mc{C}} B(x,\sigma + \epsilon)$. Defining $\mc{A}_1(s), \ldots, \mc{A}_N(s)$ for a given $s > 0$ to be
	\begin{equation*}
	\mc{A}_1(s) := B(x_1,s),~~\textrm{and}~~ \mc{A}_{j + 1}(s) := B(x_{j + 1},s) \setminus \bigcup_{i = 1}^{j} B(x_i,s) ~~\textrm{for $j = 1,\ldots,N - 1$},
	\end{equation*}
	we have that $\mc{A}_1(\sigma + \epsilon),\ldots,\mc{A}_N(\sigma + \epsilon)$ is a disjoint covering of $\mc{C}_{\sigma + \delta}$, and so $\nu(\mc{C}_{\sigma + \delta}) \leq \sum_{j = 1}^{N} \nu\bigl(\mc{A}_j(\sigma + \epsilon)\bigr)$.

@@

\subsection{Proof of Proposition~\ref{prop:density_cluster_conductance}}
\label{subsec:density_cluster_conductance}
The following Lemma lower bounds the population-level uniform conductance $\Psi_{\nu,r}(\mc{C}_{\lambda,\sigma})$.
\begin{lemma}
	\label{lem:uniform_density_cluster_conductance}
	Suppose $\mc{C}_{\lambda,\sigma}$ satisfies Assumption~\ref{asmp:embedding} with respect to some $\rho \in (0,\infty)$ and $L \in [1,\infty)$. For any $0 < r \leq \sigma \cdot \sqrt{2\pi/(d + 2)}$, it holds that
	\begin{equation}
	\label{eqn:uniform_density_cluster_conductance}
	\Psi_{\nu,r}(\mc{C}_{\lambda,\sigma}) \geq \Bigl(1 - \frac{r}{4\rho L}\Bigr) \cdot \Bigl(1 - \frac{r}{\sigma}\sqrt{\frac{d + 2}{2\pi}}\Bigr)^2 \cdot \frac{\sqrt{2\pi}}{36} \cdot \frac{r}{\rho L \sqrt{d + 2}}.
	\end{equation}
\end{lemma}
Noting that $\Psi_{\Pbb,r}(\mc{C}_{\lambda,\sigma}) \geq \Psi_{\nu,r}(\mc{C}_{\lambda,\sigma}) \cdot \lambda_{\sigma}^2/\Lambda_{\sigma}^2$, Proposition~\ref{prop:density_cluster_conductance} follows from~\eqref{eqn:uniform_density_cluster_conductance}.

@@

\subsection{Population functionals, hard case}
\label{subsec:density_cluster_hard_case}
Let $\Pbb$ be the hard case distribution over rectangular domain $\mc{X}$, defined as in~\eqref{eqn:lb_density}, and $\mc{L}$ the lower half of $\mc{X}$. Suppose $r \in (0,\sigma/2)$. Then the population normalized cut $\Phi_{\Pbb,r}(\mathcal{L})$ is lower bounded,
\begin{equation}
\label{eqn:ncut_lower_set}
\Phi_{\Pbb,r}(\mc{L}) \leq \frac{8}{3} \cdot \frac{r}{\rho}.
\end{equation}
and the population local spread $s_{\Pbb,r}(\mc{X})$ is upper bounded,
\begin{equation}
\label{eqn:local_spread_ub}
s_{\Pbb,r}(\mc{X}) \geq \frac{\pi r^2 \epsilon^2}{2 \rho\sigma}
\end{equation}

\begin{proof}[Proof of~\eqref{eqn:ncut_lower_set}.]
	Noting that $\vol_{\Pbb,r}(\mathcal{L}) = \vol_{\Pbb,r}(\mathcal{X} \setminus \mathcal{L})$, it suffices to upper bound $\cut_{\Pbb,r}(\mathcal{L})$ and lower bound $\vol_{\Pbb,r}(\mathcal{L})$. Note that for any $x = (x_1,x_2) \in \mathcal{L}$, if $x_2 \leq -r$ the ball $B(x,r)$ and the set $\mathcal{X}\setminus\mathcal{L}$ are disjoint. As a result,
	\begin{equation*}
	\cut_{\Pbb,r}(\mathcal{L}) \leq \Pbb\Bigl(\bigl\{x \in \mathcal{X}: x_2 \in (-r,0)\bigr\}\Bigr) \cdot d_{\max}(\Pbb) \leq \frac{r}{2 \rho} \cdot \frac{\pi r^2}{2 \sigma \rho}.
	\end{equation*}
	On the other hand, noting that $\deg_{\Pbb,r}(x) \geq \frac{\pi r^2}{2 \sigma \rho}$ for all $x \in \mc{C}^{(1)}$ such that $\mathrm{dist}(x, \partial \mc{C}^{(1)}) > r$, we have
	\begin{align*}
	\vol_{\Pbb,r}(\mathcal{L}) & \geq \Pbb\Bigl(\bigl\{x \in \mc{C}^{(1)} \cap \mathcal{L}: \mathrm{dist}(x, \partial \mc{C}^{(1)}) > r\bigr\}\Bigr) \cdot \frac{\pi r^2}{2 \sigma \rho} \\
	& = \frac{(\sigma - 2r)(\rho - r)}{2 \sigma \rho} \cdot \frac{\pi r^2}{2 \sigma \rho}  \\
	& \geq \frac{3}{16} \cdot \frac{\pi r^2}{2 \sigma \rho}
	\end{align*}
	where the last inequality follows since $r \leq \frac{1}{4}\sigma < \frac{1}{4}\rho$. 
\end{proof}

\begin{proof}[Proof of~\eqref{eqn:local_spread_ub}.]
	\textcolor{red}{(TODO)}
\end{proof}

@@

To prove Theorem~\ref{thm:ppr_lb}, we use Theorem~\ref{thm:normalized_cut_ppr}, Proposition~\ref{prop:sample_to_population_1} and~\eqref{eqn:ncut_lower_set} to show that the cluster estimate $\wh{C}$ must have a small normalized cut. On the other hand, in Lemma~\ref{lem:normalized_cut_lb} we establish that any set $Z \subseteq X$ which is close to $\mc{C}^{(1)}[X]$---meaning $\vol_{n,r}(Z \vartriangle \mc{C}^{(1)}[X])$ is small---has a large normalized cut.
\begin{lemma}
	\label{lem:normalized_cut_lb}
	With probability at least $1 - \exp\set{-2n\delta^2(\vol_{\Pbb,r}(\mathcal{X}))^2} - \frac{12\sigma \rho}{r^2} \exp\set{-\frac{\delta^2\epsilon r^2 n}{\rho \sigma(3 + \delta)}} - \frac{2\rho}{r}\exp\set{-\frac{\delta^2\pi r^3n}{2\sigma\rho^2(3 + \delta)}}$, the following statement holds:
	\Phi_{n,r}(Z) \geq \frac{(1 - \delta)^2}{2(1 + \delta)\pi} \left(1 - 2 \frac{\sigma \rho}{(1 - \delta) r^2 n^2} \vol_{n,r}(Z \vartriangle \mc{C}^{(1)}[X]) \right) \frac{\epsilon^2 r}{\sigma},
	for all $Z \subseteq X$.
\end{lemma}
We therefore conclude that $\vol_{n,r}(\wh{C} \vartriangle \mc{C}^{(1)}[X])$ must be large. Now, we give a formal proof Theorem~\ref{thm:ppr_lb}, and then of Lemma~\ref{lem:normalized_cut_lb}.
\begin{proof}[Proof (of Theorem~\ref{thm:ppr_lb})]
